Mechanised Modal Model Theory

### 2019-11-18 p-modalmodels/paper.tex
backup before revising

\subsubsection{Filtration}
Given \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}, the finite model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} constructed by selection is the filtration of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}} by the set of subformulas of \HOLinline{\HOLFreeVar{\ensuremath{\phi}}}. In HOL, we write \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{\ensuremath{\Sigma}}} for the filtration of the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} via the set \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}}. The formalization is almost a direct translation of the mathematical proof in \cite{Blackburn}. For the only difference: the world set of the filtrated model using the standard proof is a set of equivalence classes of worlds. To make the type of the finite model same as the model we start with, instead of taking the set of equivalence classes as the world set, we pick only one representative from each equivalence class. Hence we can prove:
\begin{thm}
{\upshape\cite[Theorem 2.41 (Finite model property, via filtration)]{Blackburn}}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\sb{\mathrm{1}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\;\HOLBoundVar{w}\sb{\mathrm{2}}.\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
\end{holmath}
\end{thm}
where \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} above is taken as \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\;(\HOLConst{subforms}\;\HOLFreeVar{\ensuremath{\phi}})}, and has the same type as \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}.

@@

\item A predicate with its symbol denoted by $p$ will hold for a list $zs$ of equivalence classes if and only if once we have a list $zr$ such that the $k$-th member is a representative of the $k$-th member of $zs$, the set of elements \HOLinline{\HOLFreeVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}} such that \HOLinline{(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLFreeVar{j}).\HOLFieldName{Pred}\;\HOLFreeVar{p}\;\HOLFreeVar{zr}} is in $U$. 
\end{itemize}
\begin{defn}
{\upshape\cite[Definition A.18 (Ultraproduct of First-Order Models)]{Blackburn}}
\begin{holmath}
  {}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLTokenDefEquality{}\\
\;\;\HOLTokenLeftrec{}\HOLFieldName{Dom}\;:=\;\HOLConst{ultraproduct}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;(\HOLConst{Doms}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}});\\
\;\;\HOLFieldName{Fun}\;:=\\
\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{n}\;\HOLBoundVar{zs}.\\
\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{y}\;\HOLTokenBar{}\\
\;\;\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Dom})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{=}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Fun}\;\HOLBoundVar{n}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U}\HOLTokenRightbrace{});\\
\;\;\HOLFieldName{Pred}\;:=\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{zs}.\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Pred}\;\HOLBoundVar{p}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})\HOLTokenRightrec{}
\end{holmath}
\end{defn}
Here we fix the representative of each equivalence class \HOLinline{\HOLFreeVar{\ensuremath{f\sb{U}}}} to be \HOLinline{\HOLConst{CHOICE}\;\HOLFreeVar{\ensuremath{f\sb{U}}}}. The function \HOLinline{\HOLConst{Doms}} takes a family of first-order models to the family of their domains.

>>>>

\subsubsection{Filtration}
Given \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}},\HOLFreeVar{w}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}}, the finite model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} constructed by selection is the filtration of \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}} by the set of subformulas of \HOLinline{\HOLFreeVar{\ensuremath{\phi}}}. In HOL, we write \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\;\HOLFreeVar{\ensuremath{\Sigma}}} for the filtration of the model \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}} via the set \HOLinline{\HOLFreeVar{\ensuremath{\Sigma}}}. The formalization is almost a direct translation of the mathematical proof in Blackburn \emph{et al}~\cite{Blackburn}. For the only difference: the world set of the filtrated model using the standard proof is a set of equivalence classes of worlds. To make the type of the finite model same as the model we start with, instead of taking the set of equivalence classes as the world set, we pick only one representative from each equivalence class. Hence we can prove:
\begin{thm}
{\upshape\cite[Theorem 2.41 (Finite model property, via filtration)]{Blackburn}}
\begin{holmath}
  \ensuremath{\HOLTokenTurnstile}(\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model}),(\HOLFreeVar{w}\sb{\mathrm{1}} :\ensuremath{\beta})\ensuremath{\Vdash}\;(\HOLFreeVar{\ensuremath{\phi}} :\ensuremath{\alpha}\;\HOLTyOp{form})\;\HOLSymConst{\HOLTokenImp{}}\\
\;\;\;\;\;\;\;\HOLSymConst{\HOLTokenExists{}}(\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}} :(\ensuremath{\alpha},\;\ensuremath{\beta})\;\HOLTyOp{model})\;(\HOLBoundVar{w}\sb{\mathrm{2}} :\ensuremath{\beta}).\;\HOLBoundVar{w}\sb{\mathrm{2}}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}},\HOLBoundVar{w}\sb{\mathrm{2}}\ensuremath{\Vdash}\;\HOLFreeVar{\ensuremath{\phi}}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLConst{FINITE}\;\HOLBoundVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}\HOLSymConst{\ensuremath{\sp{W}}}
\end{holmath}
\end{thm}
where \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{2}}} above is taken as \HOLinline{\HOLConst{FLT}\;\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}\;(\HOLConst{subforms}\;\HOLFreeVar{\ensuremath{\phi}})}, and has the same type as \HOLinline{\HOLFreeVar{\ensuremath{\mathfrak{M}}}\sb{\mathrm{1}}}.

@@

\item A predicate with its symbol denoted by $p$ will hold for a list $zs$ of equivalence classes if and only if once we have a list $zr$ such that the $k$-th member is a representative of the $k$-th member of $zs$, the set of elements \HOLinline{\HOLFreeVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}} such that \HOLinline{(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLFreeVar{j}).\HOLFieldName{Pred}\;\HOLFreeVar{p}\;\HOLFreeVar{zr}} is in $U$. 
\end{itemize}
\begin{defn}
{\upshape\cite[Definition A.18 (Ultraproduct of First-Order Models)]{Blackburn}}
\begin{holmath}
  {}^{\sf f}\Pi\sb{\HOLFreeVar{U}}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLTokenDefEquality{}\\
\;\;\HOLTokenLeftrec{}\HOLFieldName{Dom}\;:=\;\HOLConst{ultraproduct}\;\HOLFreeVar{U}\;\HOLFreeVar{J}\;(\HOLConst{Doms}\;\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}});\\
\;\;\HOLFieldName{Fun}\;:=\\
\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{n}\;\HOLBoundVar{zs}.\\
\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{y}\;\HOLTokenBar{}\\
\;\;\;\;\;\;\;\;\;\;(\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{j}.\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenImp{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Dom})\;\HOLSymConst{\HOLTokenConj{}}\\
\;\;\;\;\;\;\;\;\;\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;\HOLBoundVar{y}\;\HOLBoundVar{j}\;\HOLSymConst{=}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Fun}\;\HOLBoundVar{n}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\\
\;\;\;\;\;\;\;\;\;\;\HOLFreeVar{U}\HOLTokenRightbrace{});\\
\;\;\HOLFieldName{Pred}\;:=\\
\;\;\;\;(\HOLTokenLambda{}\HOLBoundVar{p}\;\HOLBoundVar{zs}.\;\HOLTokenLeftbrace{}\HOLBoundVar{j}\;\HOLTokenBar{}\;\HOLBoundVar{j}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{J}\;\HOLSymConst{\HOLTokenConj{}}\;(\HOLFreeVar{\ensuremath{\mathfrak{M}\sp{\sf s}}}\;\HOLBoundVar{j}).\HOLFieldName{Pred}\;\HOLBoundVar{p}\;(\HOLConst{MAP}\;(\HOLTokenLambda{}\HOLBoundVar{\ensuremath{f\sb{U}}}.\;\HOLConst{CHOICE}\;\HOLBoundVar{\ensuremath{f\sb{U}}}\;\HOLBoundVar{j})\;\HOLBoundVar{zs})\HOLTokenRightbrace{}\;\HOLSymConst{\HOLTokenIn{}}\;\HOLFreeVar{U})\HOLTokenRightrec{}
\end{holmath}
\end{defn}
Here we fix the representative of each equivalence class \HOLinline{\HOLFreeVar{\ensuremath{f\sb{U}}}} to be \HOLinline{\HOLConst{CHOICE}\;\HOLFreeVar{\ensuremath{f\sb{U}}}}. The function \HOLinline{\HOLConst{Doms}} takes a family of first-order models to the family of their domains.
